---
title: "Fig1D MOLM13"
author: "Karen Chu"
date: "1/8/2020"
output: html_document
---

Reviewer said length may be why there are more edit sites in 3'UTR. 
Normalize length somehow.

Christina says to do binomial z-score to test if the editing frequency in 3'UTR is different from CDS.
http://www.cogsci.ucsd.edu/~dgroppe/STATZ/binomial_ztest.pdf

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries.

```{r library}
library(openxlsx)
library(biomaRt)
library(stringr)
library(dplyr)
library(ggplot2)
```

Work dir.

```{r work dir}
work.dir <- "~/mount/chuk/Fig1D_MOLM13/"
output.dir <- "~/mount/chuk/Fig1D_MOLM13/output/"
figures.dir <- "~/mount/chuk/Fig1D_MOLM13/figures/"
```

Import data.
Already filtered for padj < 0.05

```{r data}
molm13 <- read.xlsx("~/mount/chuk/data_resources/molm13_snp_counts_dedupped_significant.xlsx")

molm.filter <- molm13 [ molm13$diff.frequency >= 0.1 &
                          molm13$ADA.fpkm >= 5 &
                          molm13$DCD.fpkm >= 5 &
                          molm13$MIG.fpkm >= 5, ]
```


Obtain number of A in 5'UTR.

```{r five utr}
setwd(output.dir)

# ensembl <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")
# 
# results.five.utr <- getBM(attributes = c('entrezgene_id',
#                                  '5utr'),
#                   filters = 'entrezgene_id',
#                   values = unique(molm.filter$entrez.id),
#                   mart = ensembl)
# 
# write.csv(results.five.utr, "MOLM13_HyperTRIBE_5UTR_sequences.csv", row.names = FALSE)

results.five.utr <- read.csv("MOLM13_HyperTRIBE_5UTR_sequences.csv")
```

Will result in multiple CDS called for one gene.

Obtain number of A in CDS.

```{r a cds}
setwd(output.dir)

# ensembl <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")

# results.cds <- getBM(attributes = c('entrezgene_id',
#                                  'coding'),
#                   filters = 'entrezgene_id',
#                   values = unique(molm.filter$entrez.id),
#                   mart = ensembl)
# 
# write.csv(results.cds, "MOLM13_HyperTRIBE_CDS_sequences.csv", row.names = FALSE)

results.cds <- read.csv("MOLM13_HyperTRIBE_CDS_sequences.csv")
```

Obtain number of A in 3'UTR.

```{r a cds}
setwd(output.dir)
#ensembl <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")

# results <- getBM(attributes = c('entrezgene_id',
#                                 '3utr'),
#                  filters = 'entrezgene_id',
#                  values = unique(molm.filter$entrez.id),
#                  mart = ensembl)

# write.csv(results, "MOLM13_HyperTRIBE_3UTR_sequences.csv", row.names = FALSE)

three.utr.seq <- read.csv("MOLM13_HyperTRIBE_3UTR_sequences.csv")
```

Remove Sequence Unavailable rows.

```{r remove}
results.five.utr.final <- results.five.utr [ results.five.utr$X5utr!="Sequence unavailable", ]
results.cds.final <- results.cds [ results.cds$coding!="Sequence unavailable", ]
results.three.utr.final <- three.utr.seq [ three.utr.seq$X3utr != "Sequence unavailable",]

# Subset data so CDS only contains CDS hypertribe and 3'UTR only contains 3'UTR hypertribe
# cds.entrez.id <- molm.filter [ molm.filter$annotation == "cds", ]$entrez.id
# results.cds.final <- results.cds.final [ results.cds.final$entrezgene_id %in% cds.entrez.id, ]
# 
# three.utr.entrez.id <- molm.filter [ molm.filter$annotation == "utr3", ]$entrez.id
# results.three.utr.final <- results.three.utr.final [ results.three.utr.final$entrezgene_id %in% three.utr.entrez.id, ]
```

Count total number of A's.

```{r total number of A's}
setwd(output.dir)

# Count number of A in CDS
results.five.utr.final$total.num.of.A <- str_count(results.five.utr.final$X5utr, pattern = "A")
results.cds.final$total.num.of.A <- str_count(results.cds.final$coding, pattern = "A")
results.three.utr.final$total.number.of.A <- str_count(results.three.utr.final$X3utr, pattern = "A")

# df = results.cds.final or results.three.utr.final
# region.name = cds or utr3
append.num.of.edit.sites <- function(df, region.name) {
  
  # Append column with number of edit sites for each gene.
  molm.filter.region.name <- molm.filter [ molm.filter$annotation == region.name, ]
  num.edit.sites.per.gene <- as.data.frame( table(molm.filter.region.name$entrez.id) )
  colnames(num.edit.sites.per.gene) <- c("entrezgene_id", "num.edit.sites")
  
  # Merge genes with edit site
  results.merge.subset <- merge(df, num.edit.sites.per.gene, by = "entrezgene_id")

  # Append back genes with zero edit sites
  merged.entrez.id <- results.merge.subset$entrezgene_id
  zero.edit.sites <- df [ !(df$entrezgene_id %in% merged.entrez.id), ]
  zero.edit.sites$num.edit.sites <- 0

  results.final.with.edit.sites <- rbind(results.merge.subset, zero.edit.sites)
  
  return(results.final.with.edit.sites)

}

results.five.utr.final.with.freq <- append.num.of.edit.sites(results.five.utr.final, "utr5")
results.cds.final.with.freq <- append.num.of.edit.sites(results.cds.final, "cds")
results.three.utr.final.with.freq <- append.num.of.edit.sites(results.three.utr.final, "utr3")

# write.csv(results.five.utr.final.with.freq, "results.five.utr.final.with.freq.csv", row.names = FALSE)
# write.csv(results.cds.final.with.freq, "results.cds.final.with.freq.csv", row.names = FALSE)
# write.csv(results.three.utr.final.with.freq, "results.three.utr.final.with.freq.csv", row.names = FALSE)
```

Compute z-score.

```{r compute z score}
setwd(output.dir)

# Concatenate the total number of edit sites and total number of A's
total.num.edit.sites.cds <- sum(results.cds.final.with.freq$num.edit.sites)
total.num.of.A.cds <- sum(results.cds.final.with.freq$total.num.of.A)

total.num.edit.sites.three.utr <- sum(results.three.utr.final.with.freq$num.edit.sites)
total.num.of.A.three.utr <- sum(results.three.utr.final.with.freq$total.number.of.A)

P <- total.num.edit.sites.cds / total.num.of.A.cds
C <- total.num.edit.sites.three.utr
N <- total.num.of.A.three.utr

z.score <- ( C - N*P ) / ( sqrt(N*P*(1-P)) )

output.z.score <- data.frame( CDS_total.num.edit.sites = total.num.edit.sites.cds,
                              CDS_total.num.of.A = total.num.of.A.cds,
                              ThreeUTR_total.num.edit.sites = total.num.edit.sites.three.utr,
                              ThreeUTR_total.num.of.A = total.num.of.A.three.utr,
                              P = P,
                              C = C,
                              N = N, 
                              z.score = z.score)

# write.csv(output.z.score, "output.z.score.stats.csv", row.names = FALSE)
```

Fisher exact test
columns: # A's edited vs # A's not edited
rows: 5UTR, 3UTR, CDS vs 3UTR

```{r fisher test}
# Concatenate the total number of edit sites and total number of A's
total.num.edit.sites.five.utr <- sum(results.five.utr.final.with.freq$num.edit.sites)
total.num.of.A.five.utr <- sum(results.five.utr.final.with.freq$total.num.of.A)

total.num.edit.sites.cds <- sum(results.cds.final.with.freq$num.edit.sites)
total.num.of.A.cds <- sum(results.cds.final.with.freq$total.num.of.A)

total.num.edit.sites.three.utr <- sum(results.three.utr.final.with.freq$num.edit.sites)
total.num.of.A.three.utr <- sum(results.three.utr.final.with.freq$total.number.of.A)

total.num.of.A.edited.in.all <- sum( total.num.edit.sites.five.utr, 
                                     total.num.edit.sites.cds, 
                                     total.num.edit.sites.three.utr )

total.num.of.A.NOT.EDITED.in.all <- sum( total.num.of.A.five.utr - total.num.edit.sites.five.utr,
                                         total.num.of.A.cds - total.num.edit.sites.cds,
                                         total.num.of.A.three.utr - total.num.edit.sites.three.utr )

# Prepare fisher input
fisher.input <- data.frame( num.of.A.NOT.edited = c( total.num.of.A.NOT.EDITED.in.all,
                                                     (total.num.of.A.three.utr - total.num.edit.sites.three.utr) ),
                            num.of.A.edited = c( total.num.of.A.edited.in.all, 
                                                 total.num.edit.sites.three.utr ) )
rownames(fisher.input) <- c("5UTR.3UTR.CDS", "3UTR")
fisher.input.matrix <- as.matrix(fisher.input)

# Fisher exact test
fisher.test(fisher.input.matrix, alternative="greater")
```






























