---
title: "Generate gene list for heatmap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries.

```{r libraries}
library(openxlsx)
library(dplyr)
```

Import data of significant edit sites.

```{r data}
setwd("/Users/karen/mount/chuk/LT_vs_MPPs/generate_list_of_genes_for_heatmap/")
#setwd("~/sshfs_mount/chuk/LT_vs_MPPs/generate_list_of_genes_for_heatmap/")
lt <- read.xlsx("mouse_hsc_snp_counts_dedupped_significant.xlsx", sheet=1)
st <- read.xlsx("mouse_hsc_snp_counts_dedupped_significant.xlsx", sheet=2)
mpp2 <- read.xlsx("mouse_hsc_snp_counts_dedupped_significant.xlsx", sheet=3)
mpp4 <- read.xlsx("mouse_hsc_snp_counts_dedupped_significant.xlsx", sheet=4)
```

Combine all cell types into one dataframe called "all.cell.types.no.filter"

```{r aggregate all cell types}
# Create cell.type column
lt$cell.type <- "LT"
st$cell.type <- "ST"
mpp2$cell.type <- "MPP2"
mpp4$cell.type <- "MPP4"

# Subset dataframe for necessary columns
lt.subset <- lt %>% dplyr::select(c("chr", "pos", "strand", "gene.symbol", "entrez.id", "diff.frequency", "cell.type") )
st.subset <- st %>% dplyr::select(c("chr", "pos", "strand", "gene.symbol", "entrez.id", "diff.frequency", "cell.type") )
mpp2.subset <- mpp2 %>% dplyr::select(c("chr", "pos", "strand", "gene.symbol", "entrez.id", "diff.frequency", "cell.type") )
mpp4.subset <- mpp4 %>% dplyr::select(c("chr", "pos", "strand", "gene.symbol", "entrez.id", "diff.frequency", "cell.type") )

# Merge all cell type data together.
all.cell.types.no.filter <- rbind( lt.subset, st.subset, mpp2.subset, mpp4.subset )
```

Filter by fpkm >=1 and diff.freq >=0.1.

```{r filter}
filter.function <- function(df) {

  df.filter <- df [ df$ADA.fpkm >=5 &
                    df$DCD.fpkm >=5 &
                    df$MIG.fpkm >=5 &
                    df$diff.frequency >=0.1, ]

  return(df.filter)

}

# filter.function <- function(df) {
# 
#   df.filter <- df [ df$ADA.fpkm >=1 &
#                     df$DCD.fpkm >=1 &
#                     df$MIG.fpkm >=1 &
#                     df$diff.frequency >=0.1, ]
# 
#   return(df.filter)
# 
# }

lt.filter <- filter.function(lt)
st.filter <- filter.function(st)
mpp2.filter <- filter.function(mpp2)
mpp4.filter <- filter.function(mpp4)
```

Obtain gene names after filter.

all.cell.types.no.filter.diff.freq is the dataframe with all cell types that contain genes that pass the fpkm and diff.freq filters.


```{r obtain gene names}
# Obtain genes that pass filters
genes.after.filter <- unique( c( unique(lt.filter$entrez.id), 
                                 unique(st.filter$entrez.id),
                                 unique(mpp2.filter$entrez.id),
                                 unique(mpp4.filter$entrez.id) ) )

# Filter whole list to only include genes that passed the fpkm and diff.freq filters.
all.cell.types.no.filter.subset <- all.cell.types.no.filter [ all.cell.types.no.filter$entrez.id %in% genes.after.filter, ]
all.cell.types.no.filter.diff.freq <- all.cell.types.no.filter.subset %>% dplyr::select(c("gene.symbol", "entrez.id", "diff.frequency", "cell.type"))
```


Obtain the maximum diff.frequency for each gene in each cell type.


```{r obtain max diff.freq}
# Obtain maximum diff.frequency for each gene in each cell type
obtain.max.diff.freq <- function(df, cell.name) {
  
  df.subset <- subset(df, cell.type==cell.name)
  df.max.diff.freq <- tapply(df.subset$diff.frequency, df.subset$gene.symbol, max)
  df.max.diff.freq.final <- data.frame( gene.symbol = names(df.max.diff.freq),
                                        diff.frequency = df.max.diff.freq )
  
  return(df.max.diff.freq.final)
  
}

lt.max.diff.freq <- obtain.max.diff.freq(all.cell.types.no.filter.diff.freq, "LT")
st.max.diff.freq <- obtain.max.diff.freq(all.cell.types.no.filter.diff.freq, "ST")
mpp2.max.diff.freq <- obtain.max.diff.freq(all.cell.types.no.filter.diff.freq, "MPP2")
mpp4.max.diff.freq <- obtain.max.diff.freq(all.cell.types.no.filter.diff.freq, "MPP4")

# Merge all cell types with max diff.freq by gene symbol
lt.st <- merge(lt.max.diff.freq, st.max.diff.freq, by="gene.symbol", all=TRUE)
colnames(lt.st) <- c("gene.symbol", "LT.diff.freq", "ST.diff.freq")
lt.st.mpp2 <- merge(lt.st, mpp2.max.diff.freq, by="gene.symbol", all=TRUE)
colnames(lt.st.mpp2) <- c("gene.symbol", "LT.diff.freq", "ST.diff.freq", "MPP2.diff.freq")
lt.st.mpp2.mpp4 <- merge(lt.st.mpp2, mpp4.max.diff.freq, by="gene.symbol", all=TRUE)
colnames(lt.st.mpp2.mpp4) <- c("gene.symbol", "LT.diff.freq", "ST.diff.freq", 
                               "MPP2.diff.freq", "MPP4.diff.freq")

# Convert NA to 0 
# lt.st.mpp2.mpp4 [ is.na(lt.st.mpp2.mpp4) ] <- 0
```

Export gene list with max diff.freq values.

```{r write csv}
setwd("/Users/karen/mount/chuk/LT_vs_MPPs/generate_list_of_genes_for_heatmap/")
write.csv(lt.st.mpp2.mpp4, "HSPC_fpkm_greaterthanorequalto_5_diff.freq_greaterthanorequalto_0.1_genelist_with_maximum_diff.freq_values.csv", row.names = FALSE)
```


Identify cell type specific genes by clustering
cell type specific have max diff.freq >=0.1 and all other cell types have max.diff.freq == NA.

```{r genes specific to each cell type}
setwd("/Users/karen/mount/chuk/LT_vs_MPPs/generate_list_of_genes_for_heatmap/")

max.diff.freq <- lt.st.mpp2.mpp4

lt.unique <- max.diff.freq [ max.diff.freq$LT.diff.freq >= 0.1 &
                             is.na(max.diff.freq$ST.diff.freq) &
                             is.na(max.diff.freq$MPP2.diff.freq) &
                             is.na(max.diff.freq$MPP4.diff.freq), ]

st.unique <- max.diff.freq [ max.diff.freq$ST.diff.freq >= 0.1 &
                             is.na(max.diff.freq$LT.diff.freq) &
                             is.na(max.diff.freq$MPP2.diff.freq) &
                             is.na(max.diff.freq$MPP4.diff.freq), ]

mpp2.unique <- max.diff.freq [ max.diff.freq$MPP2.diff.freq >= 0.1 &
                             is.na(max.diff.freq$LT.diff.freq) &
                             is.na(max.diff.freq$ST.diff.freq) &
                             is.na(max.diff.freq$MPP4.diff.freq), ]

mpp4.unique <- max.diff.freq [ max.diff.freq$MPP4.diff.freq >= 0.1 &
                             is.na(max.diff.freq$LT.diff.freq) &
                             is.na(max.diff.freq$ST.diff.freq) &
                             is.na(max.diff.freq$MPP2.diff.freq), ]

write.csv(lt.unique, "HSPC_fpkm_greaterthan_1_diff.freq_greaterthanorequalto_0.1_genelist_with_maximum_diff.freq_values_LT-UNIQUE.csv", row.names = FALSE)

write.csv(st.unique, "HSPC_fpkm_greaterthan_1_diff.freq_greaterthanorequalto_0.1_genelist_with_maximum_diff.freq_values_ST-UNIQUE.csv", row.names = FALSE)

write.csv(mpp2.unique, "HSPC_fpkm_greaterthan_1_diff.freq_greaterthanorequalto_0.1_genelist_with_maximum_diff.freq_values_MPP2-UNIQUE.csv", row.names = FALSE)

write.csv(mpp4.unique, "HSPC_fpkm_greaterthan_1_diff.freq_greaterthanorequalto_0.1_genelist_with_maximum_diff.freq_values_MPP4-UNIQUE.csv", row.names = FALSE)
```











