---
title: "Prepare betabinom input"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r library}
library(openxlsx)
```

Set working directory

```{r setwd}
folder <- "/Users/karen/mount/chuk/LT_vs_MPPs/"
#folder <- "/Users/chuk/sshfs_mount/chuk/LT_vs_MPPs/"
```

Import data.

Statistically significant edit sites were filtered for fpkm >= 5 and diff.freq >= 0.1, and the maximum diff.frequency in each cell type for those genes were put into a table called "HSPC_fpkm_greaterthan_5_diff.freq_greaterthanorequalto_0.1_genelist_with_maximum_diff.freq_values.csv". All the other files are just subsets of this one for LT, ST, MPP2, or MPP4-unique genes with a threshold of diff.freq >= 0.1 for the cell type of interest and diff.freq = 0 for all other cell types.

```{r import}
setwd(folder)

lt <- read.csv("generate_list_of_genes_for_heatmap/HSPC_fpkm_greaterthan_1_diff.freq_greaterthanorequalto_0.1_genelist_with_maximum_diff.freq_values_LT-UNIQUE.csv")
st <- read.csv("generate_list_of_genes_for_heatmap/HSPC_fpkm_greaterthan_1_diff.freq_greaterthanorequalto_0.1_genelist_with_maximum_diff.freq_values_ST-UNIQUE.csv")
mpp2 <- read.csv("generate_list_of_genes_for_heatmap/HSPC_fpkm_greaterthan_1_diff.freq_greaterthanorequalto_0.1_genelist_with_maximum_diff.freq_values_MPP2-UNIQUE.csv")
mpp4 <- read.csv("generate_list_of_genes_for_heatmap/HSPC_fpkm_greaterthan_1_diff.freq_greaterthanorequalto_0.1_genelist_with_maximum_diff.freq_values_MPP4-UNIQUE.csv")
```

Import SNP count data.

```{r snp}
setwd(folder)

snp.counts <- read.xlsx("snp_counts/Mouse_HSPC_snps_count.xlsx")
snp.counts <- subset(snp.counts, dbsnp=="")
```

Venn diagram of overlapping genes in cell-type-unique lists.

```{r venn diagram}
setwd(paste0(folder, "prepare_betabinom_input/"))

library("VennDiagram")

vd <- venn.diagram(
  x = list(
    "LT-unique genes" = unique(lt$gene.symbol),
    "ST-unique genes" = unique(st$gene.symbol),
    "MPP2-unique genes" = unique(mpp2$gene.symbol),
    "MPP4-unique genes" = unique(mpp4$gene.symbol)
  ),
  filename = NULL,
  col = "transparent",
  fill = c("red", "blue", "yellow", "green"),
  alpha = 0.50,
  cex = 3,
  cat.cex = 1.5
)

library(grDevices)

png(file="cell-type_specific_genes_venn.diagram.png", 800, 800 )
par(mar=c(5,5,5,5))
grid.draw(vd)
dev.off()

```


Obtain read counts for edit sites that appear in the cell-type-unique lists.

```{r obtain counts}
lt.counts <- snp.counts [ snp.counts$gene.symbol %in% lt$gene.symbol, ]
st.counts <- snp.counts [ snp.counts$gene.symbol %in% st$gene.symbol, ]
mpp2.counts <- snp.counts [ snp.counts$gene.symbol %in% mpp2$gene.symbol, ]
mpp4.counts <- snp.counts [ snp.counts$gene.symbol %in% mpp4$gene.symbol, ]

setwd(paste0(folder, "prepare_betabinom_input/"))

write.csv(lt.counts, "Mouse_HSPC_LT-unique_betabinom_input.csv")
write.csv(st.counts, "Mouse_HSPC_ST-unique_betabinom_input.csv")
write.csv(mpp2.counts, "Mouse_HSPC_MPP2-unique_betabinom_input.csv")
write.csv(mpp4.counts, "Mouse_HSPC_MPP4-unique_betabinom_input.csv")
```

Shared LT ST
Shared MPP2 MPP4

```{r shared cells}
setwd(paste0(folder, "prepare_betabinom_input/"))

shared.lt.st <- rbind(lt.counts, st.counts)
shared.mpp2.mpp4 <- rbind(mpp2.counts, mpp4.counts)

write.csv(shared.lt.st, "Mouse_HSPC_SHARED_LT_ST-unique_betabinom_input.csv")
write.csv(shared.mpp2.mpp4, "Mouse_HSPC_SHARED_MPP2_MPP4-unique_betabinom_input.csv")
```

















