---
title: "Supp Fig3I GeneExpression Independent"
author: "Karen Chu"
date: "12/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries.

```{r libraries}
library(openxlsx)
```

Set working directory.

```{r set wd}
folder <- "/Users/karen/mount/chuk/Fig3B_betabinomial_on_LSC_and_LSK_unique_targets/Supp_Fig3I_GeneExpression_independent/"
```

Import data.

```{r import}
setwd(folder)

heatmap.combined.final <- read.csv("../heatmap_combined.final.csv")

genex <- read.csv("../deseq2/LSC_ADA_vs_LSK_ADA_DESeq2Results.csv")
```

Which pass filters
p.adj >= 0.05 or p.adj < 0.05 & log2FC < 0.263 (aka FC < 1.2)

```{r filters}
setwd(folder)

first.filter.index <- which( genex$padj >= 0.05 )
second.filter.index <- which( genex$padj < 0.05 & genex$log2FoldChange < 0.263 )
filter.index <- unique( c(first.filter.index, second.filter.index) ) # combine indices of both filters
genex.independent.genes <- genex[filter.index, ] # obtain data for GE independent genes

genex.independent.genes.heatmap <- heatmap.combined.final [ heatmap.combined.final$entrez.id %in% genex.independent.genes$entrez.id, ] # subset heatmap for GE independent

write.csv(genex.independent.genes.heatmap, "genex.independent.genes.heatmap.csv", row.names = FALSE)
```

Prepare dataframe for heatmap.

```{r prep}
heatmap.combined.final <- genex.independent.genes.heatmap

# Get diff.freq for heatmap
heatmap.diff.freq.input <- heatmap.combined.final [ ,grepl("max.diff.freq", colnames(heatmap.combined.final))]

# Get edit sites
heatmap.edit.sites <- heatmap.combined.final [ ,grepl("edit.sites", colnames(heatmap.combined.final))]
heatmap.edit.sites [ is.na(heatmap.edit.sites) ] <- 0
  
# Get gene expression for heatmap
heatmap.gene.expression <- heatmap.combined.final [ ,grepl("MIG|ADA", colnames(heatmap.combined.final))]
```

Plot heatmap.

```{r heatmap}
setwd(folder)

library(ComplexHeatmap)

png("LSC_LSK_diff.freq_greaterthanorequalto_0.6_heatmap_GE_independent.png", 1000, 800)
Heatmap(heatmap.diff.freq.input,
        clustering_method_rows = "centroid",
        col=colorRamp2(c(min(heatmap.diff.freq.input), 0.1, max(heatmap.diff.freq.input)), c("blue", "white", "red")),
        column_title = "Editing Frequency",
        name="Editing Frequency",
        show_row_names = FALSE,
        cluster_columns = FALSE) +
  Heatmap(log2(heatmap.edit.sites+1),
          column_title = "Edit Sites",
          name="log2( Edit Sites +1)",
          show_row_names = FALSE,
          cluster_columns = FALSE) +
  Heatmap(heatmap.gene.expression,
          column_title = "Gene Expression",
          name="Z-transformed VST read counts",
          show_row_names = FALSE,
          cluster_columns = FALSE) 

# Add borders
decorate_heatmap_body("Editing Frequency", {

    grid.rect(gp = gpar(fill = "transparent", col = "black", lwd = 2))

})

decorate_heatmap_body("log2( Edit Sites +1)", {

    grid.rect(gp = gpar(fill = "transparent", col = "black", lwd = 2))

})

decorate_heatmap_body("Z-transformed VST read counts", {

    grid.rect(gp = gpar(fill = "transparent", col = "black", lwd = 2))

})

dev.off()
```

Manual reorder

```{r obtain order}
# We want to flip the dark red at the bottom to be at the top in LSK

h1 <- Heatmap(heatmap.diff.freq.input,
        clustering_method_rows = "centroid",
        col=colorRamp2(c(min(heatmap.diff.freq.input), 0.1, max(heatmap.diff.freq.input)), c("blue", "white", "red")),
        column_title = "Editing Frequency",
        name="Editing Frequency",
        show_row_names = FALSE,
        cluster_columns = FALSE) +
  Heatmap(log2(heatmap.edit.sites+1),
          column_title = "Edit Sites",
          name="log2( Edit Sites +1)",
          show_row_names = FALSE,
          cluster_columns = FALSE) +
  Heatmap(heatmap.gene.expression,
          column_title = "Gene Expression",
          name="Z-transformed VST read counts",
          show_row_names = FALSE,
          cluster_columns = FALSE) 

row.order <- row_order(h1)
row.order <- row.order[[1]]
heatmap.combined.final.order <- heatmap.combined.final[row.order,]
plot(heatmap.combined.final.order$LSK.A.max.diff.freq) # plot to see where the diff.freq dips
index.of.interest <- which(heatmap.combined.final.order$entrez.id==252875)

# Reverse order of the dark red to put it at the top
row.order.new <- c( rev(row.order[1:index.of.interest]), row.order[index.of.interest+1:length(row.order)] )
row.order.new <- row.order.new [ !(is.na(row.order.new)) ]
heatmap.combined.final.reorder <- heatmap.combined.final[row.order.new,]
```

Plot heatmap.

```{r heatmap prep and plot}
# Prepare for heatmap
heatmap.combined.final <- heatmap.combined.final.reorder

# Get diff.freq for heatmap
heatmap.diff.freq.input <- heatmap.combined.final [ ,grepl("max.diff.freq", colnames(heatmap.combined.final))]

# Get edit sites
heatmap.edit.sites <- heatmap.combined.final [ ,grepl("edit.sites", colnames(heatmap.combined.final))]
heatmap.edit.sites [ is.na(heatmap.edit.sites) ] <- 0
  
# Get gene expression for heatmap
heatmap.gene.expression <- heatmap.combined.final [ ,grepl("MIG|ADA", colnames(heatmap.combined.final))]

# Plot heatmap
setwd(folder)

png("LSC_LSK_diff.freq_greaterthanorequalto_0.6_heatmap_GE_independent_manual_reorder.png", 1000, 800)
Heatmap(heatmap.diff.freq.input,
        col=colorRamp2(c(min(heatmap.diff.freq.input), 0.1, max(heatmap.diff.freq.input)), c("blue", "white", "red")),
        column_title = "Editing Frequency",
        name="Editing Frequency",
        show_row_names = FALSE,
        cluster_rows = FALSE,
        cluster_columns = FALSE) +
  Heatmap(log2(heatmap.edit.sites+1),
          column_title = "Edit Sites",
          name="log2( Edit Sites +1)",
          show_row_names = FALSE,
          cluster_columns = FALSE) +
  Heatmap(heatmap.gene.expression,
          column_title = "Gene Expression",
          name="Z-transformed VST read counts",
          show_row_names = FALSE,
          cluster_columns = FALSE) 

# Add borders
decorate_heatmap_body("Editing Frequency", {

    grid.rect(gp = gpar(fill = "transparent", col = "black", lwd = 2))

})

decorate_heatmap_body("log2( Edit Sites +1)", {

    grid.rect(gp = gpar(fill = "transparent", col = "black", lwd = 2))

})

decorate_heatmap_body("Z-transformed VST read counts", {

    grid.rect(gp = gpar(fill = "transparent", col = "black", lwd = 2))

})

dev.off()
```






