---
title: "Distance of Msi2 HyperTRIBE edited sites from nearest Msi2 motif"
output: html_document
---

Purpose: Create a probability density plot of Zi-Ning's histogram, which plots the distance between Msi2 HyperTRIBE edited sites and the nearest Msi2 motif.

Method: Load the alignMotifsWorkspace.RData object Zi-Ning provided us.

Results: /figures/dist_msi2_hypertribe_edited_sites_to_nearest_msi2_motif.pdf
/figures/probability_of_Msi2_HyperTRIBE_from_Msi2_motif.txt


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load}
setwd("/Users/chuk/Documents/MSI2-hyperTRIBE/MOLM13/")

load("data/alignMotifsWorkspace.RData")
```

Below is copied from Zi-Ning's alignMotifs.R file.

```{r histogram}
# histogram
df <- data.frame(dist=query.df$dist.to.motif)
ggplot(data=df, aes(df$dist)) + 
  geom_histogram(breaks=seq(-500,500, by=20), 
                 col="white", 
                 fill="blue", 
                 alpha=0.2) + 
  labs(title="Distance to nearest motif") +
  labs(x="Distance (bp)") +
  labs(y="Count")
```

Convert histogram to PDF plot.

```{r PDF}
save.path <- "/Users/chuk/Documents/MSI2-hyperTRIBE/MOLM13/figures/"
pdf(paste0(save.path, "dist_msi2_hypertribe_edited_sites_to_nearest_msi2_motif.pdf"), 10, 10, useDingbats = F )
ggplot(df, aes(x=df$dist)) + geom_density(fill="dodgerblue") +
  scale_x_continuous(limits = c(-500, 500)) +
  labs(title="Distance to nearest motif (from Zi-Ning's Rdata)\n") +
  labs(x="\nDistance (bp)") +
  labs(y="Density\n") +
  theme_classic() +
  theme(plot.title = element_text(size=30), 
        axis.text=element_text(size=40), 
        axis.text.x = element_text(angle = 90),
        axis.title.x = element_text(size=40), 
        axis.title.y = element_text(size=40))
dev.off()
```

Calculate the probability of an Msi2 HyperTRIBE edited site to be observed within 500 bp away from an Msi2 motif.

```{r probability}
ecdf.df <- ecdf(df$dist)
ecdf.values <- data.frame(x = environment(ecdf.df)$x, y = environment(ecdf.df)$y)

output <- ""

for (i in c(1, 10, seq(0, 500, 50))) {
  
  if(i==0) next
  
  ecdf.interval.i <- subset(ecdf.values, x <= i) # +250 to negative infinity
  ecdf.interval.neg.i <- subset(ecdf.values, x < -i) # -250 to negative infinity

  probability.pos.i.to.neg.i <- (ecdf.interval.i[ nrow(ecdf.interval.i),]$y) - (ecdf.interval.neg.i[ nrow(ecdf.interval.neg.i),]$y)

  probability.output <- capture.output( cat("Probability of finding a Msi2 HyperTribe edited site ", i, "bp away from Msi2 motif in MOLM13 is ", probability.pos.i.to.neg.i, "\n\n") )
  
  output <- c(output, probability.output)
  
}

writeLines(output, paste0(save.path, "probability_of_Msi2_HyperTRIBE_from_Msi2_motif.txt"))
```











